Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-02-23 ms.tex
lx

\subsection{Nuclear X-ray Luminosities of Quiescent Galaxies}
\label{sec:mdot}

Figure \ref{fig:vwSources} shows that the CNM heating rate is on
average approximately constant with SMBH mass, except for the largest
black holes with $M_{\bullet} \gtrsim 5\times 10^{8}M_{\odot}$.  For
fixed wind heating $\dot{M}/\dot{M}_{\rm edd}$ increases with SMBH
mass $\propto M_{\bullet}^{0.5(0.8)}$ for core(cusp) galaxies,
respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]).
This appears in contradiction with the nuclear X-ray luminosities of
quiescent galaxies, which (though showing broad scatter) instead on
average obey $\langle L_X/L_{\rm edd}\rangle \sim \Mbh^\alpha$, where
$\alpha\simeq -0.25$ \citep{Pellegrini10, Miller+15}.
Figure~\ref{fig:miller} shows our predicted value of $\langle
L_{X}/L_{\rm edd} \rangle =\epsilon_X \Mdot c^2$ as a function of
black hole mass for core (blue) and cusp (green) galaxies.  The left
panel assumes a constant low radiative efficiency $\epsilon_X =
10^{-4}$ typical of those estimated for low luminosity AGN (e.g.,
\citealt{DiMatteo+00}, \citealt{Ho:2009a}) and SgrA* (e.g.,
\citealt{Quataert+99}, \citealt{Quataert:2004a}), while the right
panel shows an efficiency of $\epsilon_X= \dot{M}/\dot{M}_{\rm edd}$
that depends on the Eddington ratio in the way predicted by simple
models for radiatively inefficient accretion flows
(\citealt{Narayan&Yi95}).  Superimposed with the theoretical curves
are symbols showing X-ray measurements (red stars) and upper limits
(red triangles) from \citet{Miller+15}, along with the best power-law
fit (red line; {\bf BDM: change to this!}) and the approximate X-ray
detection threshold (gray line).  Finally, shown with purple and
orange lines are the maximum accretion rates for thermally stable
accretion (eq.~[\ref{eq:Mdotmax}]) and set by Ia blow-out
(eq.~[\ref{eq:eddr_Ia}]), as well as the maximum accretion rate
allowed by Compton heating (eq.~[\ref{eq:MdotC}]).

For either of our assumptions about radiative efficiency our predicted
trend in $\langle L_{X}/L_{\rm edd}\rangle$ for core and cusp galaxies
is opposite the observationally inferred one.  However, the observed
$\langle L_{X}/L_{\rm edd}\rangle$ do imply the accretion rates of
observed galaxies are below the maximum ones for thermally stable
accretion if $\epsilon_X \lesssim 10^{-4}$.  This suggests that most
nearby galaxies are consistent with accreting in a thermally stable
steady state, even if our estimate of their predicted heating does not
match the observed trend with $M_{\bullet}$ perfectly.  On the other
hand if $\epsilon_{X} \sim 10^{-5}$ then the trend would be consistent
with the accretion rate set by Ia blow-out.  Also note that in the
case of $\dot{M}-$dependent radiative efficiency ($\epsilon_X =
\dot{M}/\dot{M}_{\rm edd}$), the predicted branches of $\langle
L_{X}/L_{\rm edd}\rangle$ decreases with decreasing $M_{\bullet}$,
calling into question attempts to extrapolate the $\langle L_{X}
\rangle - M_{\bullet}$ relationship to low $M_{\bullet}$ in order to
constrain the occupation fraction of black holes in the lowest mass
galaxies (\citealt{Miller+15}).

>>>>

\subsection{Nuclear X-ray Luminosities}
\label{sec:mdot}

Figure \ref{fig:vwSources} shows that the CNM heating rate due to the average star formation history of a galaxy is predicted to be approximately constant with SMBH mass, except for the largest black holes with $M_{\bullet} \gtrsim 5\times 10^{8}M_{\odot}$.  For a fixed wind heating parameter, $\dot{M}/\dot{M}_{\rm edd}$ increases with SMBH
mass $\propto M_{\bullet}^{0.5(0.8)}$ for core(cusp) galaxies, respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]).  This dependence, show more precisely in Figure XXX, appears in contradiction with the nuclear X-ray luminosities of
quiescent galaxies, which (though showing several orders of magnitude scatter) instead on average obey $\langle L_X/L_{\rm edd}\rangle \sim \Mbh^\alpha$, where $\alpha \in [-0.4,-0.2]$ \citep{Zhang+09, Pellegrini10, Miller+15}.

Figure~\ref{fig:miller} shows our resulting predictions for $\langle L_{X}/L_{\rm edd} \rangle =\epsilon_X \Mdot c^2$ as a function of black hole mass for core (blue) and cusp (green) galaxies.  The left
panel is calculated assuming a constant low value for the radiative efficiency $\epsilon_X =
10^{-4}$ typical of those estimated for low luminosity AGN (e.g.,
\citealt{DiMatteo+00}, \citealt{Ho:2009a}) and SgrA* (e.g.,
\citealt{Quataert+99}, \citealt{Quataert:2004a}), while the right
panel shows an efficiency of $\epsilon_X= \dot{M}/\dot{M}_{\rm edd}$
that depends on the Eddington ratio in the way predicted by models for radiatively inefficient accretion flows (\citealt{Narayan&Yi95}).  Superimposed with the theoretical curves
are symbols showing X-ray measurements (stars) and upper limits
(triangles) compiled by \citet{Miller+15}, along with the best power-law
fit (red line) and the X-ray detection threshold (gray line).  Finally, purple, orange, and {\bf brown} lines
orange lines show the maximum accretion rates for thermally stable
accretion (eq.~[\ref{eq:Mdotmax}]), as set by SN Ia blow-out
(eq.~[\ref{eq:eddr_Ia}]), and as allowed by Compton heating feedback (eq.~[\ref{eq:MdotC}]), respectively.

To zeroth order our predicted values of $\langle L_{X}/L_{\rm edd}\rangle$ for average star formation histories overlap with the data reasonably well.  However, our predicted increase of $\langle L_{X}/L_{\rm edd}\rangle$ with black hole mass appears to be in disagreement with observed decrease of $\langle L_{X}/L_{\rm edd}\rangle$ with $M_{\bullet}$ for both core and cusp galaxies.  Better agreement is obtained by taking into account the fact that core(cusp) galaxie are more commonly associated with larger(smaller) black hole masses, such that when both samples are combined one predicts a flatter dependence of $\langle L_X/L_{\rm edd}\rangle$ of $M_{\bullet}$.   The large scatter observed in $L_{X}/L_{\rm edd}$ of typically 2-4 orders of magnitude is also natural within a steady-accretion scenario due to the strong sensitivity of the Eddington ratio on the heating rate, $\dot{M}/\dot{M}_{\rm edd} \propto v_{w}^{-3.8(-2.4)}$ for core(cusp) galaxies, respectively (Fig.~\ref{fig:mdot_mass}; eq.~[\ref{eq:eddr_analytic}]) and the likewise sensitive dependence of the heating rate on the intermittency of the star formation history (Fig.~NICK).  Alternatively, in the case of an old stellar population incapable of supporting steady accretion, resulting in cyclic accretion episodes (e.g.~\citealt{Ciotti+10}), this large scatter can be understood by the wind range of gas masses experienced over the course of a cycle.  The cyclic model also predicts that a higher fraction of $L_{X}$ nuclei with smaller stellar ages, consistent with the finding by \citet{Pellegrini10} that galaxies with younger stellar populations host less X-ray luminous nuclei.  However, this observation can also be understood in the stable accretion picture by the fact that young stars provide sufficient stellar wind heating to produce a thermally stable accretion state.

However, the observed values of 
$\langle L_{X}/L_{\rm edd}\rangle$ do imply the accretion rates of
observed galaxies are below the maximum ones for thermally stable
accretion if $\epsilon_X \lesssim 10^{-4}$.  This suggests that most
nearby galaxies are consistent with accreting in a thermally stable
steady state, even if our estimate of their predicted heating does not
match the observed trend with $M_{\bullet}$ perfectly.  On the other
hand if $\epsilon_{X} \sim 10^{-5}$ then the trend would be consistent
with the accretion rate set by Ia blow-out.  Also note that in the
case of $\dot{M}-$dependent radiative efficiency ($\epsilon_X =
\dot{M}/\dot{M}_{\rm edd}$), the predicted branches of $\langle
L_{X}/L_{\rm edd}\rangle$ decreases with decreasing $M_{\bullet}$,
calling into question attempts to extrapolate the $\langle L_{X}
\rangle - M_{\bullet}$ relationship to low $M_{\bullet}$ in order to
constrain the occupation fraction of black holes in the lowest mass
galaxies (\citealt{Miller+15}).

Past studies have found that core galaxies show a wide range of nuclear radio and X-ray luminosities, while cusp galaxies are confined below a threshold (e.g., \citealt{Bender+89, Pellegrini99, Capetti&Balmaverde05}).
